
# Loading packages and data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# Loading packages

library(pacman)
# base packages
library(conflicted)

p_load(readxl, units, janitor, plm, tidyverse, knitr, gridExtra, stargazer,  cowplot, ggplot2, ggrepel, gtsummary)

# geo packages
p_load(sf, sp, raster, rgdal, ggmap, rgeos) 

conflict_prefer_all("dplyr")
options(dplyr.summarise.inform = FALSE)
```

```{r include=FALSE}

# Get all data

 # Loading factory data
load("C:/Users/smshi/Dropbox/Research/BD-RMG-Women/data/factories.Rdata")

matched_data_01 %>%
  filter(!is.na(fac_type)) -> matched_data_01

matched_data_09 %>%
  filter(!is.na(fac_type)) -> matched_data_09

load("C:/Users/smshi/Dropbox/Research/BD-RMG-Women/data/fac_upazilas.Rdata")

# change all to integers

factory_upazilas %>%
  mutate(across(-upazila, ~as.integer(.x))) -> factory_upazilas

factory_upazilas <- factory_upazilas %>%
  mutate(across(-upazila, ~as.integer(.)))

ninetyone_sample <- factory_upazilas$ipum1991
twentyone_sample <- factory_upazilas$ipum2001
twenty11_sample <-  factory_upazilas$ipum2011

# Individual data

load("C:/Users/smshi/OneDrive/Documents/large_datasets/BD HH or Micro data/IPUMS-I/data_cleaned_indiv.Rdata")
```


# The RMG industry in Bangladesh {#rmghistory}


```{r facgrowth, fig.cap = "Expansion of RMG factories 1978:2006", fig.align='center'}

all_facs <- rbind(
  matched_data_01 %>% select(bgmea_num, date_est, fac_type),
  matched_data_09 %>% select(bgmea_num, date_est, fac_type)) %>% unique()

all_facs %>%
  group_by(date_est) %>%
  summarise(new = n()) %>%
  ggplot(aes(x = date_est, y = cumsum(new))) +
  geom_line() + 
  labs(x = "Year", y = "Number of factories",
  caption = "Data Source: Various BGMEA datasets") -> facgrowth

ggsave("figures/facgrowth.png", plot = facgrowth, dpi = 600)

rm("facgrowth", "all_facs")
```

```{r, knitandwoven, fig.cap = "Knit and Woven Exports Over the Years", fig.align='center' warning=FALSE}

matched_data_01 %>%
  filter(exist91 == 1) %>%
  group_by(fac_type) %>%
  tally() %>%
  left_join(matched_data_01 %>% filter(exist01 == 1) %>%
  group_by(fac_type) %>%
  tally(), by = "fac_type") -> knit_v_wov

matched_data_09 %>%
  filter(exist06 == 1) %>%
  group_by(fac_type) %>%
  tally() %>%
  left_join(knit_v_wov, by = "fac_type") -> knit_v_wov

names(knit_v_wov) <- c("fac_type", "2006", "1991", "2001")

knit_v_wov <- knit_v_wov %>%
  mutate(fac_type = as.character(fac_type)) %>%
  mutate(fac_type = case_when(fac_type == "0" ~ "Woven",
                              fac_type == "0.5" ~ "Mixed",
                              fac_type == "1" ~ "knit")) %>%
  gather(year, fac_num, -fac_type)
  
  
p1 <- ggplot(knit_v_wov, aes(fill = fac_type, y=fac_num, x=year)) + 
    geom_bar(position="fill", stat="identity") +
  labs(x = "Year", y = "Share of RMG Factories", fill = "", 
       caption = "Data Source: BGMEA Directory 2000-01 & 2009-110") +
  theme(legend.position="bottom")


export_data <- read_excel("C:/Users/smshi/Dropbox/Research/Data/BD RMG factory data/exports_bgmea.xlsx")%>%
  select(year, knit, woven) %>%
  pivot_longer(-year, names_to = "Type", values_to = "Exports") %>%
  filter(year <= 2011)

p2 <- ggplot(export_data, aes(x=year, y = Exports)) +
  geom_line(aes(col = Type))+
  labs(y = "Millions of USD", x = "Year", col = "",
       caption = "Data Source: BGMEA (2022)") +
  theme(legend.position="bottom")
  
grid.arrange(p1, p2, nrow =1) -> knitWov

ggsave("figures/knitWov.png", plot = knitWov, dpi = 600)

rm("export_data", "knit_v_wov", "knitWov", "p1", "p2")
```

## Reproductive Behavior

```{r ageReproduction, fig.cap = "Age specific marriage and own children in household in sample areas", fig.align='center'}

data %>%
  filter(geo3_bd1991 %in% ninetyone_sample | geo3_bd2001 %in% twentyone_sample | geo3_bd2011 %in% twenty11_sample) %>%
  filter(sex == 2 & age >= 10 & age <= 40) %>%
  group_by(age, year) %>%
  summarize(ever_married = round(100*sum(marst %in% c(2,3,4)) / n(), digits =2)) %>%
  ggplot(aes(x = age, y = ever_married, col = as.factor(year))) +
  geom_line() +
  labs(x = "Age", y = "Percent Ever Married", col = "Census Year") +
  theme(legend.position="bottom") -> p1
  
data %>%
  filter(geo3_bd1991 %in% ninetyone_sample | geo3_bd2001 %in% twentyone_sample |
           geo3_bd2011 %in% twenty11_sample) %>%
  filter(sex == 2 & age >= 10 & age <= 40) %>%
  group_by(age, year) %>%
  summarize(fert = round(mean(nchild), digits =2)) %>%
  ggplot(aes(x = age, y = fert, col = as.factor(year))) +
  geom_line() +
  labs(x = "Age", y = "Nu. Own Child in HH", col = "Census Year") +
  theme(legend.position="bottom") -> p2

grid.arrange(p1, p2, nrow =1) -> marfert

ggsave("figures/marfert.png", plot = marfert, dpi = 600)

rm("p1", "p2", "marfert")
```


# Data {#data}

## Factory and Export Data

###  Summary table FAC
```{r, summaryfactable, include=FALSE, eval=FALSE}

# Factory Statistics

matched_data_01 %>% 
  select(machine, exist91, fac_type, machine) %>% as.data.frame() %>%
  mutate(knit = case_when(fac_type == 1 ~ 1, TRUE ~0), 
         woven = case_when(fac_type == 0 ~ 1, TRUE ~ 0)) %>% select(-fac_type) %>%
  stargazer(type = "latex", omit.summary.stat = c("max", "min"), digits = 1,
            covariate.labels = c("Num. of Machines", "Existed in 1991", "Knit Factory", "Woven Factory"), 
            title = "Summary Statistics of RMG Factories in Sample", header = FALSE)
```


###  Size dist

```{r, facsize-dist, fig.cap = "Factory Size distribution in 1991 and 2001", fig.align='center', warning=FALSE}

matched_data_01 %>%
  filter(fac_type == 0) %>%
  mutate(Year = case_when(exist91 == 1 ~ "<1992", TRUE ~"1992-01")) %>%
  ggplot(aes(machine, pad = FALSE, colour = as.factor(Year))) +
  stat_ecdf(geom = "step") + 
  labs(x = "Nu. of Machines", y = "Percentage", colour = "Established on",   caption = "Woven Factories") +
  theme(legend.position="bottom") -> p1


matched_data_01 %>%
  filter(fac_type == 1) %>%
  mutate(Year = case_when(exist91 == 1 ~ "<1992", TRUE ~"1992-01")) %>%
  ggplot(aes(machine, pad = FALSE, colour = as.factor(Year))) +
  stat_ecdf(geom = "step") + 
  labs(x = "Nu. of Machines", y = "Percentage", colour = "Established on",   caption = "Knit Factories") +
  theme(legend.position="bottom") -> p2

grid.arrange(p1, p2, nrow =1) -> facSizeDist

ggsave("figures/facSizeDist.png", plot = facSizeDist, dpi = 600)

rm("p1", "p2", "facSizeDist")
```


## Population census and Labor force survey


```{r, facNOfac-census91, warning=FALSE, message=FALSE}

upazilas91  <- "C:/Users/smshi/OneDrive/Documents/large_datasets/BD HH or Micro data/IPUMS-I/geo3_bd1991/geo3_bd1991.shp"

upazilas91 <- st_read(upazilas91, quiet = TRUE) %>%
  st_as_sf(factory_upazilas91) %>% st_transform(3106) %>%
  select(contains("IP")) %>%
  clean_names() %>%
  mutate(geo3_bd1991 = as.integer(ipum1991)) %>%
  mutate(area = st_area(geometry) %>% drop_units()) %>%
  as.data.frame() %>% mutate(area = area/1000000) %>% select(geo3_bd1991, area)


data %>%
  filter(year == 1991) %>%
  left_join(upazilas91) %>%
  mutate(factory = case_when(
    geo3_bd1991 %in% ninetyone_sample ~ 1, TRUE ~ 0)) %>%
  group_by(geo3_bd1991) %>%
  mutate(pop = n()/0.1) %>%
  ungroup() %>%
  mutate(density = pop/area) %>%
  select(geo3_bd1991, density, electric, urban, age, sex, yrschool, factory) %>%
  group_by(geo3_bd1991) %>%
  summarise(across(everything(), ~mean(.,na.rm=T))) %>% 
  ungroup() -> data91 

data91 %>%
  mutate(Factory = ifelse(factory == 1, "Factory by 2006", "No Factory by 2006")) %>% select(-c(geo3_bd1991, factory)) %>%
    tbl_summary(by= Factory,
                statistic = list(everything() ~ "{mean} ({sd})"),
                missing = 'no') %>%
    add_n() %>% add_p() %>%  modify_header(label = "**Variable**") %>% # update the column header
  bold_labels() -> table_19_FacnoFac

# using the {gt} package
as_gt(table_19_FacnoFac) %>%
  gt::tab_header(title = "Table 2. Difference between Sub-districts with or Without Factory") %>% gt::as_latex() %>% as.character()
```


```{r, facCorel91, warning=FALSE, message=FALSE, eval=FALSE, include=FALSE}

matched_data_01 %>%
  left_join(factory_upazilas %>% select(ipum1991, upaz2011)) %>%
  filter(exist91 == 1 & !is.na(machine)) %>%
  group_by(ipum1991) %>%
  summarise(
    total91 = sum(machine*exist91, na.rm = T),
    total01 = sum(machine*exist01, na.rm = T),
    knit91 = sum(machine*exist91*fac_type, na.rm = T),
    knit01 = sum(machine*exist01*fac_type, na.rm = T),
    wove91 = total91 - knit91,
    wove01 = total01 - knit01) %>% ungroup %>%
  mutate(knit_share91 = knit91 / total91,
         knit_share01 = knit01 / total01,
         wove_share91 = wove91 / total91,
         wove_share01 = wove01 / total01) %>% 
    mutate(geo3_bd1991 = ipum1991) %>%
    mutate_all(~replace(., is.nan(.), 0)) -> intensity

data91 %>% 
  filter(geo3_bd1991 %in% intensity$geo3_bd1991) %>%
  left_join(intensity %>% select(geo3_bd1991, wove_share91)) %>%
  select(-c(geo3_bd1991, factory)) -> cordata


corstars <-function(x, method=c("pearson", "spearman"), removeTriangle=c("upper", "lower"),
                     result=c("none", "html", "latex")){
    #Compute correlation matrix
    require(Hmisc)
    x <- as.matrix(x)
    correlation_matrix<-rcorr(x, type=method[1])
    R <- correlation_matrix$r # Matrix of correlation coeficients
    p <- correlation_matrix$P # Matrix of p-value 
    
    ## Define notions for significance levels; spacing is important.
    mystars <- ifelse(p < .0001, "****", ifelse(p < .001, "*** ", ifelse(p < .01, "**  ", ifelse(p < .05, "*   ", "    "))))
    
    ## trunctuate the correlation matrix to two decimal
    R <- format(round(cbind(rep(-1.11, ncol(x)), R), 2))[,-1]
    
    ## build a new matrix that includes the correlations with their apropriate stars
    Rnew <- matrix(paste(R, mystars, sep=""), ncol=ncol(x))
    diag(Rnew) <- paste(diag(R), " ", sep="")
    rownames(Rnew) <- colnames(x)
    colnames(Rnew) <- paste(colnames(x), "", sep="")
    
    ## remove upper triangle of correlation matrix
    if(removeTriangle[1]=="upper"){
      Rnew <- as.matrix(Rnew)
      Rnew[upper.tri(Rnew, diag = TRUE)] <- ""
      Rnew <- as.data.frame(Rnew)
    }
    
    ## remove lower triangle of correlation matrix
    else if(removeTriangle[1]=="lower"){
      Rnew <- as.matrix(Rnew)
      Rnew[lower.tri(Rnew, diag = TRUE)] <- ""
      Rnew <- as.data.frame(Rnew)
    }
    
    ## remove last column and return the correlation matrix
    Rnew <- cbind(Rnew[1:length(Rnew)-1])
    if (result[1]=="none") return(Rnew)
    else{
      if(result[1]=="html") print(xtable(Rnew), type="html")
      else print(xtable(Rnew), type="latex") 
    }
}

corstars(cordata, result="latex")
```


# Results {#results}

## Factory locations and industry shares

```{r, results='asis'}

load("C:/Users/smshi/Dropbox/Research/BD-RMG-Women/data/fac_location_result1.RData")

reg91 <- lm(has_fac91 ~ electrification + urban_share + density_1564 + yrschool, data_fac_loc_standardized %>% filter(year == 1991))

reg91_nodensity <- lm(has_fac91 ~ electrification + urban_share + yrschool, data_fac_loc_standardized %>% filter(year == 1991))

reg01 <- lm(has_fac01 ~ electrification + urban_share + density_1564 + yrschool, data_fac_loc_standardized %>% filter(year == 2001))

reg01_nodensity <- lm(has_fac01 ~ electrification + urban_share + yrschool, data_fac_loc_standardized %>% filter(year == 2001))


# Stargazer output
stargazer(reg91, reg91_nodensity, reg01, reg01_nodensity, type = "latex", digits = 2, 
          covariate.labels = c("Electrification Rate", "Urbanization Rate", "Density (Age 15-64)", "Yrs. of Schooling"),
          dep.var.labels = c("Has Factory by 1991 = 1", "Has Factory by 2001 = 1"),
          title = "LPM Estimating Whether a Sub-district Has Factory", 
          omit = c("year"), omit.stat = c("f", "ser"), no.space=TRUE, header = FALSE)

rm(reg1, reg2, data_fac_loc_standardized)
```


```{r, results='asis'}

# knit shares
reg_knit91 <- lm(dplyr::lag(knit_share) ~
                   diff_electrification + diff_urban_share + diff_pop_1564 + diff_m_ind + diff_m_schooling, 
                 data_share_reg %>% filter(year %in% c(1991, 2001)))

reg_knit01 <- lm(dplyr::lag(knit_share) ~
                   diff_electrification + diff_urban_share + diff_pop_1564 + diff_m_ind + diff_m_schooling, 
                 data_share_reg %>% filter(year %in% c(2001, 2011)))


# Wove shares
reg_woven91 <- lm(dplyr::lag(woven_share) ~
                   diff_electrification + diff_urban_share + diff_pop_1564 + diff_m_ind + diff_m_schooling, 
                 data_share_reg %>% filter(year %in% c(1991, 2001)))

reg_woven01 <- lm(dplyr::lag(woven_share) ~
                   diff_electrification + diff_urban_share + diff_pop_1564 + diff_m_ind + diff_m_schooling, 
                 data_share_reg %>% filter(year %in% c(2001, 2011)))

# Stargazer output

stargazer(reg_knit91, reg_woven91, reg_knit01, reg_woven01, type = "latex", digits = 2, column.sep.width = "-10pt",
          covariate.labels = c("$\\Delta$ Electrification Rate", "$\\Delta$ Urbanization Rate", "$\\Delta$ Pop (Age 15-64)", "$\\Delta$ \text{Male Ind LFP}", "$\\Delta$ \text{Male Schooling (<15)}"),
          dep.var.labels = c("$\\alpha_{k}$ 91", "$\\alpha_{W}$ 91", "$\\alpha_{k}$ 01", "$\\alpha_{W}$ 01"), 
          title = "$\\alpha_{k}$ and $\\alpha_{W}$ and Changes in Sub-district Charecteristics", 
          omit = c("year"), omit.stat = c("f", "ser"), no.space=TRUE, header = FALSE)


rm(reg_knit91, reg_woven91, reg_knit01, reg_woven01, data_fac_loc_standardized)
```


```{r, regwork_flfp, cache=TRUE, results='asis'}

load("C:/Users/smshi/Dropbox/Research/BD-RMG-Women/data/final_data.RData")

reg_flfp1564 <- lm(100*dif_f_lfp1564 ~ export_exposure_pcpk + factor(year) +
  lag_electrification + lag_urban_share + 
    lag_m_yrschool1564 + lag_f_yrschool1564 + lag_density +
    + lag_age_1564_pop + lag_m_lfp1564 -1, data)

reg_flfp1529 <- lm(100*dif_f_lfp1529 ~ export_exposure_pcpk + factor(year) +
  lag_electrification + lag_urban_share + 
    lag_m_yrschool1564 + lag_f_yrschool1564 + lag_density +
    + lag_age_1564_pop + lag_m_lfp1529 -1, data)

reg_flfp1520 <- lm(100*dif_f_lfp1520 ~ export_exposure_pcpk + factor(year) +
  lag_electrification + lag_urban_share + 
    lag_m_yrschool1564 + lag_f_yrschool1564 + lag_density +
    + lag_age_1564_pop + lag_m_lfp1520 -1, data)


reg_flfp1564_ind <- lm(100*dif_f_lfp_ind1564 ~ export_exposure_pcpk + factor(year) +
  lag_electrification + lag_urban_share + 
    lag_m_yrschool1564 + lag_f_yrschool1564 + lag_density +
    + lag_age_1564_pop + lag_m_lfp_ind1564 -1, data)

reg_flfp1529_ind <- lm(100*dif_f_lfp_ind_1529 ~ export_exposure_pcpk + factor(year) +
  lag_electrification + lag_urban_share + 
    lag_m_yrschool1564 + lag_f_yrschool1564 + lag_density +
    + lag_age_1564_pop + lag_m_lfp_ind1529 -1, data)

reg_flfp1520_ind <- lm(100*dif_f_lfp_ind1520 ~ export_exposure_pcpk + factor(year) +
  lag_electrification + lag_urban_share + 
    lag_m_yrschool1564 + lag_f_yrschool1564 + lag_density +
    + lag_age_1564_pop + lag_m_lfp_ind1520 -1, data)


stargazer(reg_flfp1564, reg_flfp1564_ind, reg_flfp1529, reg_flfp1529_ind,
          reg_flfp1520, reg_flfp1520_ind, type = "latex", digits = 2, 
          column.labels = c("FLFP", "FLFP-Ind", "FLFP", "FLFP-Ind", "FLFP", "FLFP-Ind"),
          covariate.labels = c("$\\Delta$ Export Exposure", "Electrification Rate", "Urbanization Rate", "Yrs. School (15-64, M)", "Yrs. School (15-64, F)", "Density", "Pop (15-64)", "$Y_{s,t-1}^{M}$"),
          dep.var.labels = c(" Ages 15-64", "Ages 15-29", "Ages 15-20"), 
          title = "Export Exposure and Female Labor Force Participation", 
          omit = c("year", "Constant"), omit.stat = c("f", "ser"), no.space=TRUE, header = FALSE)
```


```{r, regwork_marFert, cache=TRUE, results='asis'}
load("C:/Users/smshi/Dropbox/Research/BD-RMG-Women/data/final_data.RData")

reg_mar1520 <- lm(100*dif_married_1520 ~ export_exposure_pcpk + factor(year) +
  lag_electrification + lag_urban_share + 
    lag_m_yrschool1564 + lag_f_yrschool1564 + lag_density -1, data)

reg_mar2130 <- lm(100*dif_married_2130 ~ export_exposure_pcpk + factor(year) +
  lag_electrification + lag_urban_share + 
    lag_m_yrschool1564 + lag_f_yrschool1564 + lag_density -1, data)

reg_fert1520 <- lm(100*dif_fertility_1520 ~ export_exposure_pcpk + factor(year) +
  lag_electrification + lag_urban_share + 
    lag_m_yrschool1564 + lag_f_yrschool1564 + lag_density -1, data)

reg_fert2130 <- lm(100*dif_fertility_2130 ~ export_exposure_pcpk + factor(year) +
  lag_electrification + lag_urban_share + 
    lag_m_yrschool1564 + lag_f_yrschool1564 + lag_density -1, data)

reg_fert3040 <- lm(100*dif_fertility_3040 ~ export_exposure_pcpk + factor(year) +
  lag_electrification + lag_urban_share + 
    lag_m_yrschool1564 + lag_f_yrschool1564 + lag_density -1, data)




stargazer(reg_mar1520, reg_mar2130, reg_fert1520, reg_fert2130,
          reg_fert3040, type = "text", digits = 2, 
          dep.var.labels = c("Marriage Rate", "Fertility"),
          covariate.labels = c("$\\Delta$ Export Exposure", "Electrification Rate", "Urbanization Rate", "Yrs. School (15-64, M)", "Yrs. School (15-64, F)", "Density"),
          column.labels = c("Ages 15-20", "Ages 21-30", "Ages 15-20", "Ages 21-30", "Ages 30-40"), 
          title = "Export Exposure and Marriage and Age specific Fertility", 
          omit = c("year", "Constant"), omit.stat = c("f", "ser"), no.space=TRUE, header = FALSE)
```


```{r, regwork_educ, cache=TRUE, results='asis'}
load("C:/Users/smshi/Dropbox/Research/BD-RMG-Women/data/final_data.RData")

reg_enrol0509 <- lm(100*dif_f_enrollment_0509 ~ export_exposure_pcpk + factor(year) +
  lag_electrification + lag_urban_share + 
    lag_m_yrschool1564 + lag_f_yrschool1564 + lag_density + I(100*lag_m_enrollment_0509) -1, data)

reg_enrol1013 <- lm(100*dif_f_enrollment_1013 ~ export_exposure_pcpk + factor(year) +
  lag_electrification + lag_urban_share + 
    lag_m_yrschool1564 + lag_f_yrschool1564 + lag_density + I(100*lag_m_enrollment_1013) -1, data)

reg_enrol1419 <- lm(100*dif_f_enrollment_1419 ~ export_exposure_pcpk + factor(year) +
  lag_electrification + lag_urban_share + 
    lag_m_yrschool1564 + lag_f_yrschool1564 + lag_density + I(100*lag_m_enrollment_1419) -1, data)



reg_lit1419 <- lm(100*dif_f_literacy_1419 ~ export_exposure_pcpk + factor(year) +
  lag_electrification + lag_urban_share + 
    lag_m_yrschool1564 + lag_f_yrschool1564 + lag_density + I(100*lag_m_literacy_1419) -1, data)

reg_yrs1419 <- lm(dif_f_yrschool_1419 ~ export_exposure_pcpk + factor(year) +
  lag_electrification + lag_urban_share + 
    lag_m_yrschool1564 + lag_f_yrschool1564 + lag_density + lag_m_yrschool_1419
-1, data)


stargazer(reg_enrol0509, reg_enrol1013, reg_enrol1419, reg_lit1419,
          reg_yrs1419, type = "text", digits = 2, 
          dep.var.labels = c("Enrollment Rate", "Literacy", "Yrs. School"),
          covariate.labels = c("$\\Delta$ Export Exposure", "Electrification Rate", "Urbanization Rate", "Yrs. School (15-64, M)", "Yrs. School (15-64, F)", "Density", "$Y_{s,t-1}^{M}$"),
          column.labels = c("Ages 05-09", "Ages 10-13", "Ages 14-19", "Ages 14-19", "Ages 14-19"), 
          title = "Export Exposure and Human Capital Accumulation", 
          omit = c("year", "Constant"), omit.stat = c("f", "ser"), no.space=TRUE, header = FALSE)
```



\newpage

# References

::: {#refs}
:::

\newpage

# Appendix 1: Evolution of women's work and fertility in Bangladesh {#appendix1 .unnumbered}


```{r, flfp-graph, fig.cap = "Female Labor Force Participation in Bangladesh"}

flfp <- read.csv("C:/Users/smshi/Dropbox/Research/BD-RMG-Women/data/flfp.csv", check.names=FALSE) %>%
    pivot_longer(-year, names_to="country", values_to="flfp")

flfp_start <- flfp %>%
  filter(year == 1990)

flfp_end <- flfp %>%
  filter(year == 2019)

ggplot(flfp, aes(x = year, y = flfp)) +
  geom_line(aes(col = country)) +
  geom_line(data = subset(flfp, country =="Bangladesh"), color = "red", size =1.25) +
geom_text_repel(
    aes(label = round(flfp,1)), data = flfp_start,
    fontface ="plain", color = "black", size = 2
    ) +
  geom_text_repel(
    aes(label = round(flfp,1)), data = flfp_end,
    fontface ="plain", color = "black", size = 2
    ) +
   labs(y = "FLFP Rate (Age 15-64)", col = "Geographical Unit", caption = "Data Source: World Bank (2021)")+ theme(legend.position="bottom") -> flfP_transition

ggsave("figures/flfp_transition.png", plot = flfP_transition, dpi=600, width = 10)

rm("flfp", "flfp_start", "flfp_end", "flfP_transition")

```

```{r, fertility-graph, fig.cap = "Fertility in Bangladesh", echo=FALSE}

fertility <- read.csv("C:/Users/smshi/Dropbox/Research/BD-RMG-Women/data/fertility.csv", check.names=FALSE) %>%
    pivot_longer(-year, names_to="country", values_to="fertility")

fertility_start <- fertility %>%
  filter(year == 1971)

fertility_end <- fertility %>%
  filter(year == 2018)

ggplot(fertility, aes(x = year, y = fertility)) +
  geom_line(aes(col = country)) +
  geom_line(data = subset(fertility, country =="Bangladesh"), color = "red", size =1.25) + geom_text_repel(
    aes(label = round(fertility,1)), data = fertility_start,
    fontface ="plain", color = "black", size = 2
    ) +
  geom_text_repel(
    aes(label = round(fertility,1)), data = fertility_end,
    fontface ="plain", color = "black", size = 2
    ) +
   labs(y = "Total Fertility Rates", col = "Geographical Units",
        caption = "Data Source: World Bank (2021)") + theme(legend.position="bottom") -> fertility_transition

ggsave("figures/fertility_transition.png", plot = fertility_transition, dpi = 600, width = 10)

rm(fertility, fertility_start, fertility_end, fertility_transition)
```

\newpage 


```{r, regwork_flfp, cache=TRUE, results='asis'}

load("C:/Users/smshi/Dropbox/Research/BD-RMG-Women/data/final_data.RData")

reg_flfp1 <- lm(100*dif_f_lfp1564 ~ export_exposure_pcpk -1, data)

reg_flfp2 <- lm(100*dif_f_lfp1564 ~ export_exposure_pcpk + factor(year) -1, data)


reg_flfp3 <- lm(100*dif_f_lfp1564 ~ export_exposure_pcpk + factor(year) +
  lag_electrification + lag_urban_share -1, data)


reg_flfp4 <- lm(100*dif_f_lfp1564 ~ export_exposure_pcpk + factor(year) +
  lag_electrification + lag_urban_share + 
    lag_m_yrschool1564 + lag_f_yrschool1564 + lag_density -1, data)

reg_flfp5 <- lm(100*dif_f_lfp1564 ~ export_exposure_pcpk + factor(year) +
  lag_electrification + lag_urban_share + 
    lag_m_yrschool1564 + lag_f_yrschool1564 + lag_density +
    + lag_age_1564_pop + lag_m_lfp1564 -1, data)


stargazer(reg_flfp1, reg_flfp2, reg_flfp3, reg_flfp4, reg_flfp5, type = "latex", digits = 2, 
          column.labels = c("FLFP", "FLFP-Ind", "FLFP", "FLFP-Ind", "FLFP", "FLFP-Ind"),
          covariate.labels = c("$\\Delta$ Export Exposure", "Electrification Rate", "Urbanization Rate", "Yrs. School (15-64, M)", "Yrs. School (15-64, F)", "Density", "Pop (15-64)", "$Y_{s,t-1}^{M}$"),
          dep.var.labels = c(" Ages 15-64", "Ages 15-29", "Ages 15-20"), 
          title = "Export Exposure and Female Labor Force Participation", 
          omit = c("year", "Constant"), omit.stat = c("f", "ser"), no.space=TRUE, header = FALSE)
```

# Appendix 2: Spread of the RMG industry in Bangladesh {#appendix2 .unnumbered}

```{r, facspread, fig.cap = "Spatial spread of RMG factories in Bangladesh", cache = TRUE, results='asis', eval=FALSE, echo=FALSE}

# Loading data


# Factory Map

factory_upazilas91  <- "C:/Users/smshi/OneDrive/Documents/large_datasets/BD HH or Micro data/IPUMS-I/geo3_bd1991/geo3_bd1991.shp"

factory_upazilas91 <- st_read(factory_upazilas91, quiet = TRUE)

# Convert to Sf and transform to Mercator Gulshan
factory_upazilas91 <- st_as_sf(factory_upazilas91) %>% st_transform(3106) %>%
  select(ADMIN_NAME, contains("UP"), PARENT) %>%
  clean_names() %>%
  mutate_if(is.character,tolower)

# Bangladesh zones 0 - country map, 1 - div, 2- district, 3-sub-district, 4 - smaller
geo_bd0 <- "C:/Users/smshi/Dropbox/University of Oregon/Research/Bangladesh FLFPR/Data/Bd admin and roads/bgd_adm_bbs_20201113_shp/bgd_admbnda_adm0_bbs_20201113.shp"

bd_map0 <- st_read(geo_bd0, quiet = TRUE)
bd_map0 <- bd_map0 %>% st_transform(3106)

p1 <- ggplot(bd_map0) +
  geom_sf() +
  geom_sf(data = filter(factory, date_est<=1999), color = "black") +
  geom_sf(data = filter(factory, date_est>1999 & date_est<=2000), color = "red") +
  theme_map() + labs(title="Factories betwen 99-00")

p2 <- ggplot(bd_map0) +
  geom_sf() +
  geom_sf(data = filter(factory, date_est<=2000), color = "black") +
  geom_sf(data = filter(factory, date_est > 2000 & date_est <= 2004), color = "red")+
  theme_map() + labs(title="Factories betwen 00-04")

p3 <- ggplot(bd_map0) +
  geom_sf() +
  geom_sf(data = filter(factory, date_est<=2004), color = "black") +
  geom_sf(data = filter(factory, date_est > 2004 & date_est <= 2007), color = "red")+
  theme_map() + labs(title="Factories betwen 04-07")

p4 <- ggplot(bd_map0) +
  geom_sf() +
  geom_sf(data = filter(factory, date_est<=2007), color = "black") +
  geom_sf(data = filter(factory, date_est > 2007 & date_est <= 2011), color = "red")+
  theme_map() + labs(title="Factories betwen 07-11")

p5 <- ggplot(bd_map0) +
  geom_sf() +
  geom_sf(data = filter(factory, date_est<=2011), color = "black") +
  geom_sf(data = filter(factory, date_est > 2011 & date_est <= 2014), color = "red")+
  theme_map()+labs(title="Factories betwen 11-14")

p6 <- ggplot(bd_map0) +
  geom_sf() +
  geom_sf(data = filter(factory, date_est<=2014), color = "black") +
  geom_sf(data = filter(factory, date_est > 2014 & date_est <= 2018), color = "red")+
  theme_map() + labs(title="Factories betwen 14-18")

grid.arrange(p1, p2, p3, p4, p5, p6, nrow =3)
```

\newpage

# Appendix 3: Summary Statistics of Outcomes, Census 2001 {#appendix3 .unnumbered}